---
title: "Latent Factors of Language Disturbance and Relationships to Quantitative Speech Features: NLP + Network Analysis"
author: "Tang, Hänsel, Cong, Nikzad, Mehta, Cho, Berretta, Behbehani, Pradhan, John, & Liberman"
date: "8/5/2022"
output:
  html_document:
    toc: true
    theme: united
---


```{r setup, include=FALSE}
### Load Packages
library(effsize) #cohen d, etc.
library(lm.beta) #standardized beta
library(psych) #psych stats package, inlcudes factor analysis
library(gridExtra) #arranges ggplot grobs
library(Hmisc) #works with ggplot and others
library(ggpubr) #pairwise and statistics
library(bestNormalize) #yeojohnson and other transformations
library(mice) #multiple imputation
library(plyr) #data manipulations
library(MatchIt) #simpler matching
library(naniar) #na functions
library(arsenal) #tableby
library(hexbin) #Binning and plotting functions for hexagonal bins - network
library(networkD3) #Creates D3 JavaScript network, tree, dendrogram graphs from R.
library(tidytext) #Text mining
library(repr)#String and binary representations - Network
library(RColorBrewer)
library(ggthemes) #Preset themes for ggplot
library(tidyverse) #dplyr, ggplot, stringr, et al
library(regclass)
require(igraph)

set.seed(123)

### Load Clinical Data
load("crossdx_all.R")
load("vars_ratings.R")


### Data Dictionary
#crossdx_all - merged clinical (subset of clin vars) and NLP features
#ssd_all - just for SSD group

```

## Summary
- Objective - visualize/examine connections between clinical ratings + NLP features
- Strategy - look in cross-diagnostic sample first, then in SSD
- Features / Variables at our disposal (see colnames)
```{r}
#colnames(crossdx.all)
```

## Identifying Features

### NLP - determined using VIF
- grouped by feature category: typographical/speech quantity, graph, parts of speech, speech errors, coherence measures, lexical features, sentiment, acoustics - voice quality, acoustics - pause/speaking rate
- cutoff = 5, using tlc_sum score - similar results with using factor scores - Exception is coherence measures where all VIF were above 5, so used 10 instead
```{r}
# Typographical / speech quantity
VIF(lm(tlc_sum ~ sentence_count + word_count + mean_sentence_length + ttr, data = crossdx.all, na.action = na.omit))
cols.quantity <- c("ttr", "mean_sentence_length")

# Graph features
VIF(lm(tlc_sum ~ seq_nn + seq_ne + seq_ad + seq_awd + seq_gd + seq_aspl + seq_cc + seq_lq + seq_lscc + seq_diameter + seq_triangles + cooc_nn + cooc_ne + cooc_ad + cooc_awd + cooc_gd + cooc_aspl + cooc_cc + cooc_lq + cooc_diameter + ap_nn + ap_ne + ap_ad + ap_awd + ap_gd + ap_aspl + ap_cc + ap_lq + ap_lscc + ap_diameter + ap_triangles, data = crossdx.all, na.action = na.omit))
cols.graph <- c("seq_cc", "seq_lq", "ap_cc")

# Parts of speech
VIF(lm(tlc_sum ~ sp.pos_ADJ + sp.pos_ADP + sp.pos_ADV + sp.pos_AUX + sp.pos_CCONJ + sp.pos_DET + sp.pos_INTJ + sp.pos_NOUN + sp.pos_PART + sp.pos_PRON + sp.pos_PROPN + sp.pos_PUNCT + sp.pos_SCONJ + sp.pos_VERB, data = crossdx.all, na.action = na.omit))
cols.pos <- c("sp.pos_ADJ", "sp.pos_ADP", "sp.pos_CCONJ", "sp.pos_DET", "sp.pos_PART", "sp.pos_SCONJ")

# Speech error and dysfluencies
VIF(lm(tlc_sum ~ partial_error + repetition_error + umuh, data = crossdx.all, na.action = na.omit))
cols.error <- c("partial_error", "repetition_error", "umuh")

# Coherence Measures
VIF(lm(tlc_sum ~ mean_glove + mean_lsa + mean_w2v + tf_idf_glove + tf_idf_lsa + tf_idf_w2v, data = crossdx.all, na.action = na.omit))
cols.cohere <- "mean_lsa"

# Lexical features
VIF(lm(tlc_sum ~ lex_AoA + lex_Prevalence + lex_SemD, data = crossdx.all, na.action = na.omit))
cols.lex <- "lex_AoA"

# Sentiment
VIF(lm(tlc_sum ~ sent_valpos + sent_valneg + sent_val + sent_arousal + sent_dominance, data = crossdx.all, na.action = na.omit))
cols.sent <- c("sent_valpos", "sent_valneg", "sent_arousal")

# Acoustics - voice quality
VIF(lm(tlc_sum ~ acou_f0_mean + acou_f0_sd + acou_jitter_mean + acou_jitter_sd + acou_shimmer_mean + acou_shimmer_sd, data = crossdx.all, na.action = na.omit))
cols.acquality <- c("acou_f0_mean", "acou_f0_sd", "acou_shimmer_sd")

# Acoustics - tempo / pause
VIF(lm(tlc_sum ~ acou_pauselenth_mean + acou_pauselength_sd + acou_speakrate_mean + acou_speakrate_sd + acou_speakrate_min + acou_speakrate_max + acou_turnlatency_mean, data = crossdx.all, na.action = na.omit))
cols.actempo <- c("acou_pauselenth_mean", "acou_pauselength_sd", "acou_speakrate_mean", "acou_speakrate_min", "acou_turnlatency_mean")

```

### Clinical Categories, based on theory
```{r}
# Speech
cols.speech <- vars_ratings
cols.factors <- c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress")

# Clinical Scales
cols.negative <- c("sans_global_affect", "sans_global_alogia", "sans_global_avolition", "sans_global_asocanhed")
cols.bprs <- c("bprs_factor_anxdep", "bprs_factor_hostsus", "bprs_factor_thought", "bprs_factor_withret")
cols.functioning <- c("qls_inv_role", "qls_inv_interpersonal" )

# Cognition
cols.cog <- c("hint_inv_total", "er_inv_cr", "letter_inv_total", "category_inv_total")

```

### Plot Setup
```{r warning=FALSE}
# to play around with
P.VAL.THRESH = 0.05
RHO.THRESH = 0.3

# SSD vs. Psy spectrum
ssd.all = crossdx.all %>% filter(group == "SSD")
psy.all = crossdx.all %>% filter(group == "SSD" | group == "OPD")

# Create a manual node list:
nodes = data.frame(matrix(ncol=2, nrow=0))

categories = c("cols.factors", "cols.bprs", "cols.negative", "cols.functioning", "cols.cog",   "cols.quantity", "cols.graph", "cols.pos", "cols.error", "cols.cohere", "cols.lex", "cols.sent", "cols.acquality", "cols.actempo")

for (c in categories){
    terms = get(c)
    temp = data.frame(node = terms, category = rep(c, length(terms)))
    temp$node = as.character(temp$node)
    
    nodes = rbind(nodes, temp)
}

nodes$type = "clinical"
nodes$type[nodes$category %in% c("cols.quantity", "cols.graph", "cols.pos", "cols.error", "cols.cohere", "cols.lex", "cols.sent", "cols.acquality", "cols.actempo")] = "speech"
nodes$type[nodes$category %in% c("cols.factors")] = "factor"
nodes$type = as.factor(nodes$type)

# manually extract all edges
n_nodes = nrow(nodes)
n_edges = (n_nodes * (n_nodes - 1)) / 2
edges = data.frame(matrix(ncol=14, nrow = n_edges))

counter = 1

# create correlation edge list
for (i in 1:(nrow(nodes)-1)){
    node_1 = as.character(nodes[i, "node"])
    for (k in (i+1):nrow(nodes)){
        node_2 = as.character(nodes[k, "node"])
        
        corr = cor.test(crossdx.all[[node_1]], crossdx.all[[node_2]], method = "spearman")
        p.val = corr$p.value
        rho = corr$estimate
        rho.abs = abs(rho)
        rho.sign = if (rho >= 0) 1 else 2
        
        corr.ssd = cor.test(ssd.all[[node_1]], ssd.all[[node_2]], method = "spearman")
        p.val.ssd = corr.ssd$p.value
        rho.ssd = corr.ssd$estimate
        rho.ssd.abs = abs(corr.ssd$estimate)
        rho.ssd.sign = if (rho.ssd >= 0) 1 else 2
        
        corr.psy = cor.test(psy.all[[node_1]], psy.all[[node_2]], method = "spearman")
        p.val.psy = corr.psy$p.value
        rho.psy = corr.psy$estimate
        rho.psy.abs = abs(corr.psy$estimate)
        rho.psy.sign = if (rho.psy >= 0) 1 else 2

        edges[counter, 1] = node_1
        edges[counter, 2] = node_2
        edges[counter, 3] = rho
        edges[counter, 4] = rho.abs
        edges[counter, 5] = rho.sign
        edges[counter, 6] = p.val
        edges[counter, 7] = rho.ssd
        edges[counter, 8] = rho.ssd.abs
        edges[counter, 9] = rho.ssd.sign
        edges[counter, 10] = p.val.ssd
        edges[counter, 11] = rho.psy
        edges[counter, 12] = rho.psy.abs
        edges[counter, 13] = rho.psy.sign
        edges[counter, 14] = p.val.psy
        
        counter = counter + 1

    }
}
edges = edges %>% dplyr::rename(
                source = X1,
                target = X2,
                rho = X3,
                rho.abs = X4,
                rho.sign = X5,
                p.val = X6,
                rho.ssd = X7,
                rho.ssd.abs = X8,
                rho.ssd.sign = X9,
                p.val.ssd = X10,
                rho.psy = X11,
                rho.psy.abs = X12,
                rho.psy.sign = X13,
                p.val.psy = X14)

# filter:
edges = edges[complete.cases(edges),]
l = c(edges$source, edges$target)
nodes = nodes[nodes$node %in% l,]

nodes$category = factor(nodes$category, levels= categories)
nodes$category_num = as.numeric(nodes$category)

nodes$node = as.factor(nodes$node)

edges$source = factor(edges$source, levels = levels(nodes$node))
edges$target = factor(edges$target, levels = levels(nodes$node))

# Re-naming
nodes$category = recode(nodes$category,
       cols.factors = "Speech Factors",
       cols.acquality = "(VQ) Acoustics - Voice Quality",
       cols.actempo = "(TEM) Acoustics - Tempo",
       cols.bprs = "(BPRS) General Symptoms",
       cols.negative = "(SANS) Negative Symptoms",
       cols.functioning = "(QLS) Functional Impairment",
       cols.cog = "(COG) Cognitive Impairment",
       cols.quantity = "(QUAN) Speech Quantity", 
       cols.error = "(ERR) Speech Errors",
       cols.graph = "(GPH) Graph Measures",
       cols.pos = "(POS) Parts of Speech",
       cols.cohere = "(COS) Cosine Embedding Distances",
       cols.lex = "(LEX) Lexical Features",
       cols.sent = "(SENT) Sentiment"
       )

nodes$nodelabel = recode(nodes$node,
                    bprs_factor_anxdep = "BPRS:AnxDep",
                    bprs_factor_hostsus = "BPRS:HostSusp",
                    bprs_factor_thought = "BPRS:ThoughtDist",
                    bprs_factor_withret = "BPRS:WithdrawRet",
                    sans_global_affect = "SANS:AffectFlat",
                    sans_global_alogia = "SANS:Alogia",
                    sans_global_avolition = "SANS:Avolition",
                    sans_global_asocanhed = "SANS:Anhedonia",
                    qls_inv_role = "QLS:RoleFx",
                    qls_inv_interpersonal = "QLS:SocialFx",
                    hint_inv_total = "COG:TOM",
                    er_inv_cr = "COG:EmoProc",
                    letter_inv_total = "COG:LetterFl",
                    category_inv_total = "COG:CategoryFl",
                    tlc_3f_inefficient = "Inefficient Sp.",
                    tlc_3f_incoherent = "Incoherent Sp.",
                    tlc_3f_impexpress = "Imp. Expressivity",
                    ttr = "QUAN:TypeTokenRat",
                    mean_sentence_length = "QUAN:UtterLength",
                    seq_cc = "GPH:SeqClustCoeff",
                    seq_lq = "GPH:SeqLargestClique",
                    ap_cc = "GPH:ActionPredClustCoeff",
                    sp.pos_ADJ = "POS:Adjectives",
                    sp.pos_ADP = "POS:Adpositions",
                    sp.pos_CCONJ = "POS:CoorConjunc",
                    sp.pos_SCONJ = "POS:SubConjunc",
                    sp.pos_DET = "POS:Determiners",
                    sp.pos_PART = "POS:Particles",
                    partial_error = "ERR:PartialWords",
                    repetition_error = "ERR:RepeatedWords",
                    umuh = "ERR:FilledPauses",
                    mean_lsa = "COS:LSAcosineDist",
                    lex_AoA = "LEX:AgeofAquisition",
                    sent_valpos = "SENT:PositiveValence",
                    sent_valneg = "SENT:NegativeValence",
                    sent_arousal = "SENT:Arousal",
                    acou_f0_mean = "VQ:MeanPitch",
                    acou_f0_sd = "VQ:PitchVari",
                    acou_shimmer_sd = "VQ:ShimmerVari",
                    acou_pauselenth_mean = "TEM:MeanPauseLength",
                    acou_pauselength_sd = "TEM:PauseLengthVari",
                    acou_speakrate_mean = "TEM:MeanSpeakRate",
                    acou_speakrate_min = "TEM:MinSpeakRate",
                    acou_turnlatency_mean = "TEM:MeanTurnLatency"
                    )

# for consistent mapping of colors to categories:
color_map = c(
  "Speech Factors" = "orangered", 
  "(BPRS) General Symptoms" = "navajowhite2", 
  "(SANS) Negative Symptoms" = "moccasin", 
  "(QLS) Functional Impairment" = "mistyrose1", 
  "(COG) Cognitive Impairment" = "rosybrown1", 
  ######
  "(QUAN) Speech Quantity" = "plum2", 
  "(GPH) Graph Measures" = "thistle",
  "(POS) Parts of Speech" = "lavender", 
  "(ERR) Speech Errors" = "lightsteelblue", 
  "(COS) Cosine Embedding Distances" = "lightskyblue1", 
  "(LEX) Lexical Features" = "paleturquoise", 
  "(SENT) Sentiment" = "darkslategray1",
  "(VQ) Acoustics - Voice Quality" = "darkseagreen1", 
  "(TEM) Acoustics - Tempo" = "palegreen1")

node.categories = as.character(unique(nodes$category))
```


## Net graphs

### Plotting - cross-diagnostic

#### Speech
- current parameters (8/8/22): p=0.05, r = 0.2, seed = 122, l <- norm_coords(l, xmin=-3, xmax=2.8, ymin=-1.7, ymax=1.7), png("fig.net_crossdx_nlp3.png", 1300, 700)
```{r}
set.seed(122)

P.VAL.THRESH = 0.05
RHO.THRESH = 0.20

edges.all = edges[edges$p.val <= P.VAL.THRESH,]
edges.all = edges.all[edges.all$rho.abs >= RHO.THRESH,]

nodes.filt = filter(nodes, type %in% c("speech", "factor"))
edges.filt = filter(edges.all, source %in% nodes.filt$node & target %in% nodes.filt$node)

net.cross.speech = graph_from_data_frame(edges.filt, vertices = nodes.filt,directed = F)

pal.edge = c("lightsteelblue4", "coral3")
shapes = c("square", "circle")

V(net.cross.speech)$color = color_map[V(net.cross.speech)$category]
V(net.cross.speech)$frame.color = NA

E(net.cross.speech)$width = E(net.cross.speech)$rho.abs 
E(net.cross.speech)$curved <- 0.2
E(net.cross.speech)$color = pal.edge[E(net.cross.speech)$rho.sign]

V(net.cross.speech)$size <- scales::rescale(log(degree(net.cross.speech)), to = c(4, 13))

l <- layout_with_fr(net.cross.speech) #Resets the layout
l <- norm_coords(l, xmin=-3, xmax=2.8, ymin=-1.7, ymax=1.7)

png("fig.net_crossdx_nlp3.png", 1300, 700)

plot(net.cross.speech,
     vertex.label.cex = 1, 
     vertex.label = nodes.filt$nodelabel,
     vertex.label.color= "black",
     vertex.label.font=2,
     rescale = F, layout = l * 1.3, xlim = c(-2,2), ylim = c(-2,2))

# legend("topleft",bty = "n",
#        legend=node.categories,
#        fill=color_map[node.categories],
#        cex = 1.3)
dev.off()

```

##### Graph Features (Crossdx - speech)
- Density = 0.21
- Degree (Ineff = 8, Incoherent = 6, Impaired Ex = 7)
- Centrality (Ineff = 0.54, incohere = 0.46, imp exp = 0.51)
- Betweenness (Ineff=25.3, Incohere = 13.0, imp exp = 18.3)

```{r}
# Density
print("=== Density ===")
edge_density(net.cross.speech, loops = F)

# Degree for each node
print("=== Degree of all nodes ===")
degree(net.cross.speech, mode = "all") %>% sort(decreasing = TRUE)

# Centrality (normalized)
print("=== Normalized Centrality ===")
temp <- centr_clo(net.cross.speech, mode="all", normalized=T)$res
names(temp) <- names(closeness(net.cross.speech, mode="all", weights=NA))
temp %>% sort(decreasing = TRUE)

# Betweenness
print("=== Betweenness ===")
betweenness(net.cross.speech, directed=T, weights=NA) %>% sort(decreasing = TRUE) #Same as normalized values

# Community Detection
ceb <- cluster_edge_betweenness(net.cross.speech) 
dendPlot(ceb, mode="hclust")

# Top correlations
print("=== Top Correlations ===")
temp <- select(crossdx.all, all_of(nodes$node))
cor(temp, use = "complete.obs")[which(nodes$node == "tlc_3f_inefficient"),] %>% sort()
cor(temp, use = "complete.obs")[which(nodes$node == "tlc_3f_incoherent"),] %>% sort()
cor(temp, use = "complete.obs")[which(nodes$node == "tlc_3f_impexpress"),] %>% sort()
```



### Plotting: Just SSD

#### Speech
- Current parameters (8/8/22): p = 0.05, r=0.20, set.seed(123), l <- norm_coords(l, xmin=-3, xmax=2.8, ymin=-1.6, ymax=2.8), png("fig.net_ssd_nlp1.png", 1300, 700)
```{r}
options(repr.plot.width=22, repr.plot.height=22)
set.seed(123)

P.VAL.THRESH = 0.05
RHO.THRESH = 0.2

edges.ssd = edges[edges$p.val.ssd <= P.VAL.THRESH,]
edges.ssd = edges.ssd[edges.ssd$rho.ssd.abs >= RHO.THRESH,]

nodes.filt = filter(nodes, type %in% c("speech", "factor"))
edges.filt = filter(edges.ssd, source %in% nodes.filt$node & target %in% nodes.filt$node)

net.ssd.speech = graph_from_data_frame(edges.filt, vertices = nodes.filt, directed = F)

pal.edge = c("lightsteelblue4", "coral3")
shapes = c("square", "circle")

V(net.ssd.speech)$color = color_map[V(net.ssd.speech)$category]
V(net.ssd.speech)$frame.color = NA

E(net.ssd.speech)$width = E(net.ssd.speech)$rho.abs 
E(net.ssd.speech)$curved <- 0.2
E(net.ssd.speech)$color = pal.edge[E(net.ssd.speech)$rho.sign]

V(net.ssd.speech)$size <- scales::rescale(log(degree(net.ssd.speech)), to = c(4, 13))

l <- layout_with_fr(net.ssd.speech) #Resets the layout

l <- norm_coords(l, xmin=-3, xmax=2.8, ymin=-1.6, ymax=2.8)

png("fig.net_ssd_nlp1.png", 1300, 700)

plot(net.ssd.speech,
     vertex.label.cex = 1, 
     vertex.label = nodes.filt$nodelabel,
     vertex.label.color= "black",
     vertex.label.font=2,
     rescale = F, layout = l * 1.3, xlim = c(-2,2), ylim = c(-2,2))

dev.off()

```

##### Graph Features (SSD - speech)
- Density = 0.18
- Degree (Ineff = 9, Incoherent = 4, Impaired Ex = 2)
- Centrality (Ineff = 0.60, incohere = 0.47, imp exp = 0.37)
- Betweenness (Ineff=47.9, Incohere = 2.4, imp exp = 0.5)

```{r}
# Density
print("=== Density ===")
edge_density(net.ssd.speech, loops = F)

# Degree for each node
print("=== Degree of all nodes ===")
degree(net.ssd.speech, mode = "all") %>% sort(decreasing = TRUE)

# Centrality (normalized)
print("=== Normalized Centrality ===")
temp <- centr_clo(net.ssd.speech, mode="all", normalized=T)$res
names(temp) <- names(closeness(net.ssd.speech, mode="all", weights=NA))
temp %>% sort(decreasing = TRUE)

# Betweenness
print("=== Betweenness ===")
betweenness(net.ssd.speech, directed=T, weights=NA) %>% sort(decreasing = TRUE) #Same as normalized values

# Community Detection
ceb <- cluster_edge_betweenness(net.ssd.speech) 

dendPlot(ceb, mode="hclust")
```

