---
title: "TLC Factor Scores - Expanded Version"
author: "Sunny"
date: "12/27/2021"
output:
  html_document:
    toc: true
    theme: united
---


```{r setup, include=FALSE}
### Load Packages
library(effsize) #cohen d, etc.
library(lm.beta) #standardized beta
library(DescTools) #Fisher r to z tranformation
library(nlme) #mixed effects
#library(irr) # interrater correlation
library(psych) #psych stats package, inlcudes factor analysis
library(gridExtra) #arranges ggplot grobs
library(Hmisc) #works with ggplot and others
library(ggpubr) #pairwise and statistics
library(ggthemes) #ggplot convenient themes
library(car) #companion to applied regression
library(bestNormalize) #yeojohnson and other transformations
library(mice) #multiple imputation
library(plyr) #data manipulations
#library(Matching) #propensity score matching
#library(MatchIt) #simpler matching
library(robustbase) #adjusted box plot
library(performance) #check_outliers for mahalanobis distance, influence
library(naniar) #na functions
library(lme4) #expands on nlme with generalized linear regressions
library(glmnet) #lasso and elastic net
library(arsenal) #tableby
library(caret) #ML methods
library(lavaan) #cFA, mediation analysis
library(tidyverse) #dplyr, ggplot, stringr, et al

set.seed(123)

### Related Scripts ###
# 1) Script Compile Factors
# 2) RMD Factors Expanded ** THIS
# 3) Script Compile Networks
# 4) RMD Network V#

### Loading Data
load("crossdx_clin.R")
load("vars_ratings.R")

#Filter just SSD
ssd.clin <- filter(crossdx.clin, group == "SSD")

# Filter OPD OR SSD
opd.clin <- filter(crossdx.clin, group == "OPD" | group == "SSD")

### Data Dictionary
#XXX.tlc - just tlc + 2 sans items

```


## Summary
- N = 343 total, N=91 for just SSD
- Purpose: answer the question - what dimensions of speech disturbance should we be modeling with NLP? eFA for TLC (+2 sans items). Using FA and not PCA because interested in latent factors.
- eFA Results - 3 factor is best. Inefficient vs. incoherent vs. impaired expressivity factors. - CFA in cross diagnostic vs. SSD groups confirms eFA results.


## Sample
```{r, results='asis'}
table(crossdx.clin$group)

table(crossdx.clin$study)

tableby(group ~ demo_age + demo_gender + demo_sex + demo_race + demo_ethnicity + demo_education + demo_caregiver_education + dxcurr_psyscale + dxcurr_cat_anx + dxcurr_cat_mood + dxcurr_cat_substance + tlc_global + tlc_sum + wrat_std,
        data=crossdx.clin, numeric.stats=c("mean", "range", "sd")) %>% summary(digits = 2, digits.p=3)
```


## TLC Prevalance
- Which variables are being rated?
- Cross diagnostic - Positive in <10% of sample: clanging, neologism, word approximation, echolalia, stilted speech, blocking, self reference
- SSD: Clanging, word approximation, self-reference
```{r, include=FALSE}
### Cross diagnostic
# Print tables for each item. If >= 309 of rating 0, then >90% with 0 rating

vars_items <- vars_ratings[c(1:18, 21:22)]
save(vars_items, file = "vars_items.R")

for (i in vars_items) {
      print(i)
      print(table(crossdx.clin[[i]]))
}

### Just SSD
# 90% is 82
for (i in vars_items) {
      print(i)
      print(table(ssd.clin[[i]]))
}
```


## First-pass, all variables (n = 20) - assess requirements for FA

### Correlation Matrix:
- Cross-dx - NO correlations above 0.3: stilted speech, echolalia, self-reference
- SSD: echolalia + self reference + stilted speech
```{r}
### Cross Diagnostic
#Isolating just the items
crossdx.tlc <- select(crossdx.clin, all_of(vars_items))

#Correlation Matrix on all tlc
cor.crossdx.tlc <- cor(crossdx.tlc, use = "complete.obs")
#round(cor.crossdx.tlc, 2)

#Checking if correlations >0.3 (omitting self-correlations)
temp <- abs(cor.crossdx.tlc) >0.3 & cor.crossdx.tlc <1
rowSums(temp)

### SSD
#Isolating just the items
ssd.tlc <- select(ssd.clin, all_of(vars_items))

#Correlation Matrix on all tlc
cor.ssd.tlc <- cor(ssd.tlc, use = "complete.obs")
round(cor.ssd.tlc, 2)

#Checking if correlations >0.3 (omitting self-correlations)
temp <- abs(cor.ssd.tlc) >0.3 & cor.ssd.tlc <1
rowSums(temp)

```


### Bartlett Test of Sphericity:
- Want <0.05 to indicate significantly different from 0
- Indeed, p<<0.0001 so we're good
```{r, echo=TRUE}
# Cross-dx
cortest.bartlett(crossdx.tlc)

#SSD
cortest.bartlett(ssd.tlc)
```


### Kaiser-Meyer-Olkin Measure of Sampling Adequacy
- "MSA" Measure of sampling adequacy
- We want overall value >0.8
- Crossdx: Lowest: echolalia 0.48 - next <0.7 poverty of speech, word approx, blocking, self-reference, vocal inflection
- SSD: <0.5: echolalia + blocking + self-reference; <0.6 vocal inflection

```{r, echo=TRUE}
# Cross-dx
KMO(cor.crossdx.tlc)

#SSD
KMO(cor.ssd.tlc)
```


### Determinant
- To check for multicolinearity, want >0.00001
- cross dx 0.00007 - OK
- ssd 0.00003 - OK

```{r, echo=TRUE}
#Cross dx
det(cor.crossdx.tlc)

#SSD
det(cor.ssd.tlc)
```


### Scree plot / Kaiser Criterion
- 3 factors for both
```{r, echo=TRUE}
#Cross dx
scree(crossdx.tlc)

#SSD
scree(ssd.tlc)
```




## Second-pass, vars n = 14
- MINUS echolalia, blocking, clanging, neologism, word approximation, self-reference and stilted :: Due to low KMO, low prevalence

### Correlation Matrix:
```{r}
#New vars
vars_fa <- vars_items[-c(9, 11, 15:18)]
#vars_fa <- vars_items[-c(15:18)]

### Cross Diagnostic
#Isolating just the items
crossdx.tlc <- select(crossdx.clin, all_of(vars_fa))

#Correlation Matrix on all tlc
cor.crossdx.tlc <- cor(crossdx.tlc, use = "complete.obs")
#round(cor.crossdx.tlc, 2)

#Checking if correlations >0.3 (omitting self-correlations)
temp <- abs(cor.crossdx.tlc) >0.3 & cor.crossdx.tlc <1
rowSums(temp)

### SSD- only
#Isolating just the items
ssd.tlc <- select(ssd.clin, all_of(vars_fa))

#Correlation Matrix on all tlc
cor.ssd.tlc <- cor(ssd.tlc, use = "complete.obs")
round(cor.ssd.tlc, 2)

#Checking if correlations >0.3 (omitting self-correlations)
temp <- abs(cor.ssd.tlc) >0.3 & cor.ssd.tlc <1
rowSums(temp)

### Dataframes for the other groups for cFA below
#OPD
opd.tlc <- select(opd.clin, all_of(vars_fa))
summary(opd.tlc)

```


### Bartlett Test of Sphericity:
- Want <0.05 to indicate significantly different from 0
- Indeed, p<<0.0001 so we're good
```{r, echo=TRUE}
# Cross-dx
cortest.bartlett(crossdx.tlc)

#SSD
cortest.bartlett(ssd.tlc)
```


### Kaiser-Meyer-Olkin Measure of Sampling Adequacy
- "MSA" Measure of sampling adequacy
- We want overall value >0.8: crossdx = 0.9, ssd = 0.85
```{r, echo=TRUE}
# Cross-dx
KMO(cor.crossdx.tlc)

#SSD
KMO(cor.ssd.tlc)
```


### Determinant
- To check for multicolinearity, want >0.00001
- cross dx 0.0004 - good
- ssd 0.0004 - good

```{r, echo=TRUE}
#Cross dx
det(cor.crossdx.tlc)

#SSD
det(cor.ssd.tlc)
```


### Scree plot / Kaiser Criterion
- 2-3 factors for both
```{r, echo=TRUE}
#Cross dx
scree(crossdx.tlc)

#SSD
scree(ssd.tlc)
```


### eFA
- factoring method = principal axis
- rotation = promax (same result as oblimin)
- 3-factor a favorite
```{r, echo=TRUE}
#Cross-dx - 2 factor
model01 <- fa(crossdx.tlc, nfactors = 2, fm="pa", rotate = "promax")
fa.diagram(model01)
model01$loadings
#model01

#Cross-dx - 3-factor
model02 <- fa(crossdx.tlc, nfactors = 3, fm="pa", rotate = "promax")
fa.diagram(model02)
model02$loadings

#SSD - 2 factor
model03 <- fa(ssd.tlc, nfactors = 2, fm="pa", rotate = "promax")
fa.diagram(model03)
#model03

#SSD - 3 factor
model04 <- fa(ssd.tlc, nfactors = 3, fm="pa", rotate = "promax")
fa.diagram(model04)
model04$loadings
#model03
```


### Communality
- % Variance explained by retained factors, want this to be higher
- Not bad. I think?
```{r}
#Cross - 2 factor
model01$communality
#Cross - 3 factor
model02$communality
#SSD - 2 factor
model03$communality
```


### Eigenvalues and %variance explained
```{r}
# Extracting Rotated Eigenvalues
model01$values[1:2]
model02$values[1:3]
model03$values[1:2]

# % Variance explained
model01$values[1:2]/length(model01$values)
model02$values[1:3]/length(model02$values)
model03$values[1:2]/length(model03$values)

```


## Confirmatory Factor Analysis
- Testing 2, 3 (omit cross-loaded variables), and 4 (keep cross-loaded variables as separate factor) factor solutions. 
- Key statistics: (should report standardized statistics)
- Chi square - ideally p>0.05, but this rarely happens with larger samples
- CFI - 0 to 1, 1 being perfect. Above 0.9 or 0.95 is pretty good.
- TLI highly correlated w TLI but can be >1
- RMSEA - root mean square error of approximation is an absolute measure of fit. <= 0.05 is a close fit. 0.05 - 0.08 reasonable approximation. >-0.10 poor fit.

### 2-Factor cFA models
- 2-Factor model is poor fit in ALL samples (cross diagnostic, SSD only, psychosis spectrum)
```{r}
#2 - factor
cmod2f <- 'f1 =~ tlc_06derail + tlc_02povcontent + tlc_05tangent + tlc_13lossgoal + tlc_08illogic + tlc_12circum + tlc_14persev + tlc_07incoh + tlc_03pressure + tlc_04distract + tlc_10neologism
f2 =~ tlc_01povspeech + sans_6_vocalinf + sans_11_latency'

#Default - test in cross-dx
cmod2f_crossdx <- cfa(cmod2f, data = crossdx.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod2f_crossdx, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 2-factor from crossdx in SSD
cmod2f_ssd <- cfa(cmod2f, data = ssd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod2f_ssd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 2-factor from crossdx in OPD
cmod2f_opd <- cfa(cmod2f, data = opd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod2f_opd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

```

### 3-Factor cFA without cross-loadings
- 3-factor model without the cross loaded factors is very good for crossdx, borderline for SSD, and good for psychosis spectrum
- Fixing variance to 1 performs better than fixing the loading of first variable
```{r}
#3 - factor
cmod3f <- 'f1 =~ tlc_05tangent + tlc_12circum + tlc_03pressure + tlc_04distract + tlc_14persev
f2 =~ tlc_07incoh + tlc_08illogic + tlc_10neologism
f3 =~ tlc_01povspeech + sans_6_vocalinf + sans_11_latency'

#Default - test in cross-dx
cmod3f_crossdx <- cfa(cmod3f, data = crossdx.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod3f_crossdx, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 3-factor from crossdx in SSD
cmod3f_ssd <- cfa(cmod3f, data = ssd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod3f_ssd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 3-factor from crossdx in OPD (-distractibility)
cmod3f_opd <- cfa(cmod3f, data = opd.clin, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod3f_opd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

```

### 4-Factor Model- with separate cross-loaded factor
- Fixing variance performs better than fixing loading of first variable
- Performance is good for crossdx, acceptable for SSD, good for psychosis spectrum group
```{r}
###Fourfactor model
#4 - factor
cmod4f <- 'f1 =~ tlc_05tangent + tlc_12circum + tlc_03pressure + tlc_04distract + tlc_14persev
f2 =~ tlc_07incoh + tlc_08illogic + tlc_10neologism
f3 =~ tlc_01povspeech + sans_6_vocalinf + sans_11_latency
f4 =~ tlc_06derail + tlc_13lossgoal + tlc_02povcontent'

#Cross-dx
cmod4f_crossdx <- cfa(cmod4f, data = crossdx.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod4f_crossdx, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 4-factor in SSD
cmod4f_ssd <- cfa(cmod4f, data = ssd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod4f_ssd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#OPD
cmod4f_opd <- cfa(cmod4f, data = opd.clin, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod4f_opd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)
```

## Generate Factor Scores
- Based on eFA results because these make more sense to me.
- Grice, James W.,2001, Computing and evaluating factor scores, Psychological Methods, 01,4, 430-450. (note the typo in equation 8)
- ten Berge, Jos M.F., Wim P. Krijnen, Tom Wansbeek and Alexander Shapiro (1999) Some new results on correlation-preserving factor scores prediction methods. Linear Algebra and its Applications, 289, 311-318. - Using Ten Berge method because it's able to preserve oblique-ness of factors

```{r, include=FALSE}
### 3-Factor Solution - generating factor scores
just3factortlc <- factor.scores(crossdx.tlc, model02, method = "tenBerge", impute="mean")$scores %>% as.data.frame()
str(just3factortlc)
names(just3factortlc) <- c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress")

#Get rid of any old variables from prior runs
crossdx.clin <- select(crossdx.clin, -starts_with("tlc_3f"))

# Combine with full dataset
crossdx.clin <- cbind(crossdx.clin, just3factortlc)

### filling in missing group variables 
crossdx.clin$grid[which(is.na(crossdx.clin$group))] #all good

save(crossdx.clin, file="crossdx_clin.R")
write.csv(crossdx.clin, file="crossdx_clin.csv")

### Saving Model
save(model02, file="TLC3fmodel.R")
model02$loadings

save(model01, file="TLC2fmodel.R")

```


## Plotting Factor relationships

### Group
```{r}
# Long format for plotting together (Factors** UP TO HERE)
temp <- pivot_longer(crossdx.clin, cols = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), names_to = "factor")
justlongfactors <- temp %>% select(group, factor, value)
justlongfactors$factor <- factor(justlongfactors$factor, levels = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), labels = c("Inefficient Speech", "Incoherent Speech", "Impaired Expressivity"))

# Make group consistent with poster labels
justlongfactors$group <- as.character(justlongfactors$group)
justlongfactors$group[which(justlongfactors$group == "HC")] <- "HV"
justlongfactors$group[which(justlongfactors$group == "HSC")] <- "OD"
justlongfactors$group[which(justlongfactors$group == "OPD")] <- "PSY"
justlongfactors$group <- as.factor(justlongfactors$group)
table(justlongfactors$group)

#PLOT
ggplot(justlongfactors, aes(x = group, y=value, fill = group)) +
  geom_boxplot() + 
  facet_grid(.~factor) + 
  theme_few() + 
  scale_fill_viridis_d() + 
  labs(x = NULL, y = "Factor Z-Scores", fill = NULL) + 
  theme(text=element_text(size=24, family="Times New Roman")) +
  stat_compare_means(label = "p.signif", method = "t.test", comparisons = list( c("HV", "OD"), c("OD", "PSY"), c("PSY", "SSD"), c("HV", "PSY"), c("OD", "SSD"), c("HV", "SSD") ))

```

### Demographics (in SSD)
```{r}
ssd.clin <- filter(crossdx.clin, group == "SSD")

fig.ssd.data <- pivot_longer(ssd.clin, cols = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), names_to = "factor")
fig.ssd.data$factor <- factor(fig.ssd.data$factor, levels = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), labels = c("Inefficient Speech", "Incoherent Speech", "Impaired Expressivity"))

### AGE
ggplot(fig.ssd.data, aes(x = demo_age, y=value, color = factor)) +
  geom_point() + 
  geom_smooth(method = lm, alpha = 0.2) +
  theme_few() + 
  scale_color_brewer(palette = 'Set2') + 
  labs(x = "Age (years)" , y = "Factor Z-Scores", color = "Factors", title = "(1) Relationships to Age") + 
  theme(text=element_text(size=20, family="Times New Roman"))

lm(tlc_3f_inefficient ~ demo_age, data = ssd.clin) %>% summary()
lm.beta(lm(tlc_3f_inefficient ~ demo_age, data = ssd.clin))
lm(tlc_3f_incoherent ~ demo_age, data = ssd.clin) %>% summary()
lm.beta(lm(tlc_3f_incoherent ~ demo_age, data = ssd.clin))
lm(tlc_3f_impexpress ~ demo_age, data = ssd.clin) %>% summary()
lm.beta(lm(tlc_3f_impexpress ~ demo_age, data = ssd.clin))

### Gender
ggplot(fig.ssd.data[which(fig.ssd.data$demo_gender!="Unknown"),], aes(x = demo_gender, y=value, fill = demo_gender)) +
  geom_boxplot() + 
  facet_grid(.~factor) + 
  theme_few() + 
  scale_fill_brewer(palette = 'Accent') + 
  labs(x = NULL, y = "Factor Z-Scores", fill = NULL, title = "(2) Relationships to Gender") + 
  theme(text=element_text(size=20, family="Times New Roman")) +
  stat_compare_means(label = "p.signif", method = "t.test", comparisons = list( c("Man", "Non-Binary"), c("Non-Binary", "Woman"), c("Man", "Woman") ))

lm(tlc_3f_impexpress ~ demo_gender, data = ssd.clin) %>% summary()
lm(tlc_3f_incoherent ~ demo_gender, data = ssd.clin) %>% summary()
lm(tlc_3f_inefficient ~ demo_gender, data = ssd.clin) %>% summary()

lm(tlc_3f_impexpress ~ demo_gender + bprs_factor_hostsus , data = ssd.clin) %>% summary()
lm(tlc_3f_incoherent ~ demo_gender + bprs_factor_hostsus , data = ssd.clin) %>% summary()
lm(tlc_3f_inefficient ~ demo_gender + bprs_factor_hostsus , data = ssd.clin) %>% summary()

```


## Samples
```{r, results='asis'}

# Cross-diagnostic
tableby(group ~ study + demo_age + demo_sex + demo_gender + demo_race + demo_ethnicity + tlc_global, data = crossdx.clin) %>% summary(digits = 1, digits.pct = 1)

```


