---
title: "Latent Factors of Language Disturbance and Relationships to Quantitative Speech Features: Factor Analysis"
author: "Tang, Hänsel, Cong, Nikzad, Mehta, Cho, Berretta, Behbehani, Pradhan, John, & Liberman"
date: "7/25/2022"
output:
  html_document:
    toc: true
    theme: united
---


```{r setup, include=FALSE}
###### Project Setup ######

### Load Packages
library(psych) #psych stats package, includes factor analysis
library(plyr) #data manipulations
library(naniar) #na functions
library(arsenal) #tableby
library(GPArotation) # For Factor analysis
library(lavaan) #cFA
library(MatchIt) #Matching
library(optmatch) #Matching
library(ggthemes) #ggplot themes
library(ggpubr) #graph stuff
library(tidyverse) #dplyr, ggplot, stringr, et al


### Setting a seed for reproducibility
set.seed(123)

### Related Scripts ###
# 1) Script Compile Factors
# 2) RMD Factors Expanded ** THIS
# 3) Script Compile Networks
# 4) RMD Network V#
# 5) Script Correlations

### Loading Required Data
#load("crossdx_clin.R")
load("crossdx_clin.R")
load("vars_ratings.R")

#Filter just SSD
ssd.clin <- filter(crossdx.clin, group == "SSD")

# Filter OPD OR SSD
opd.clin <- filter(crossdx.clin, group == "OPD" | group == "SSD")

### Dataframe Dictionary
#XXX.tlc - just tlc + 2 sans items

```


## Summary
- N = 343 total cross-diagnostic dataset, N=90 for just SSD, N=137 for PS (OPD + SSD)
- This script includes exploratory + confirmatory factor analyses for the crossdiagnostic group, the SSD only group, and the psychosis spectrum group (SSD + OPD) as well as group comparisons for the factor scores.
- Purpose: answer the question - what dimensions of speech disturbance should we be modeling with NLP? eFA for TLC (+2 sans items). Using FA and not PCA because interested in latent factors.
- eFA Results - 3 factor is best. Inefficient vs. incoherent vs. impaired expressivity factors. - CFA in cross diagnostic vs. SSD groups confirms eFA results.


## Sample Characteristics
```{r, results='asis'}
table(crossdx.clin$study)

tableby(group ~ demo_age + demo_gender + demo_sex + demo_race + demo_ethnicity + demo_education + demo_caregiver_education + dxcurr_psyscale + dxcurr_cat_anx + dxcurr_cat_mood + dxcurr_cat_substance + tlc_global + tlc_sum + wrat_std,
        data=crossdx.clin, numeric.stats=c("mean", "range", "sd")) %>% summary(digits = 2, digits.p=3)
```


## TLC items: Prevalance & Sampling Adequacy

### Item Prevalence
- Cross diagnostic - Positive in <10% of sample: clanging, neologism, word approximation, echolalia, stilted speech, blocking, self reference
- SSD: Clanging, word approximation, self-reference
```{r, include=FALSE}
### Cross diagnostic
# Print tables for each item. If >= 309 of rating 0, then >90% with 0 rating

vars_items <- vars_ratings[c(1:18, 21:22)]
save(vars_items, file = "vars_items.R")

for (i in vars_items) {
      print(i)
      print(table(crossdx.clin[[i]]))
}

### Just SSD
# 90% is 82
for (i in vars_items) {
      print(i)
      print(table(ssd.clin[[i]]))
}
```

### Correlation Matrix:
- Cross-dx - NO correlations above 0.3: stilted speech, echolalia, self-reference
- SSD: echolalia + self reference + stilted speech
```{r}
### Cross Diagnostic
#Isolating just the items
crossdx.tlc <- select(crossdx.clin, all_of(vars_items))

#Correlation Matrix on all tlc
cor.crossdx.tlc <- cor(crossdx.tlc, use = "complete.obs")
#round(cor.crossdx.tlc, 2)

#Checking if correlations >0.3 (omitting self-correlations)
temp <- abs(cor.crossdx.tlc) >0.3 & cor.crossdx.tlc <1
rowSums(temp)
```

### Kaiser-Meyer-Olkin Measure of Sampling Adequacy
- "MSA" Measure of sampling adequacy
- We want overall value >0.8
- Crossdx: Lowest: echolalia 0.48 - next <0.7 poverty of speech, word approx, blocking, self-reference, vocal inflection
```{r, echo=TRUE}
# Cross-dx
KMO(cor.crossdx.tlc)

```



## eFA, vars n = 14
- EXCLUDING: echolalia, blocking, clanging, word approximation, self-reference and stilted :: Due to low KMO, low prevalence

### Correlation Matrix:
```{r}
#New vars
vars_fa <- vars_items[-c(9, 11, 15:18)]

#Isolating just the items
crossdx.tlc <- select(crossdx.clin, all_of(vars_fa))
ssd.tlc <- select(ssd.clin, all_of(vars_fa))
opd.tlc <- select(opd.clin, all_of(vars_fa))

#Correlation Matrix on all tlc
cor.crossdx.tlc <- cor(crossdx.tlc, use = "complete.obs")
#round(cor.crossdx.tlc, 2)

#Checking if correlations >0.3 (omitting self-correlations)
temp <- abs(cor.crossdx.tlc) >0.3 & cor.crossdx.tlc <1
rowSums(temp)

```

### Bartlett Test of Sphericity:
- Want <0.05 to indicate significantly different from 0
- Indeed, p<<0.0001 so we're good
```{r, echo=TRUE}
# Cross-dx
cortest.bartlett(crossdx.tlc)
```

### Kaiser-Meyer-Olkin Measure of Sampling Adequacy
- "MSA" Measure of sampling adequacy
- We want overall value >0.8: crossdx = 0.9, ssd = 0.85
```{r, echo=TRUE}
# Cross-dx
KMO(cor.crossdx.tlc)
```


### Determinant
- To check for multicolinearity, want >0.00001
- cross dx 0.0004 - good
- ssd 0.0004 - good
```{r, echo=TRUE}
#Cross dx
det(cor.crossdx.tlc)
```


### Scree plot / Kaiser Criterion
- 2-3 factors 
```{r, echo=TRUE}
#Cross dx
scree(crossdx.tlc)
```


### eFA in cross-diagnostic dataset
- factoring method = principal axis
- rotation = promax (oblimin produces consistent results)
```{r, echo=TRUE}
#Cross-dx - 2 factor
model01 <- fa(crossdx.tlc, nfactors = 2, fm="pa", rotate = "promax")
fa.diagram(model01)
model01$loadings
#model01

#Cross-dx - 3-factor
model02 <- fa(crossdx.tlc, nfactors = 3, fm="pa", rotate = "promax")
fa.diagram(model02)
model02$loadings
```


### Communality
- % Variance explained by retained factors
```{r}
#Cross - 2 factor
model01$communality
#Cross - 3 factor
model02$communality
```

### Eigenvalues and %variance explained
```{r}
# Extracting Rotated Eigenvalues
model01$values[1:2]
model02$values[1:3]

# % Variance explained
model01$values[1:2]/length(model01$values)
model02$values[1:3]/length(model02$values)
```


## Confirmatory Factor Analysis
- Testing 2, 3 (omit cross-loaded variables), and 4 (keep cross-loaded variables as separate factor) factor solutions. 
- Key statistics: TLI/CFI
- Chi square - ideally p>0.05, but this rarely happens with larger samples
- CFI - 0 to 1, 1 being perfect. Above 0.9 or 0.95 is pretty good.
- TLI highly correlated w TLI but can be >1
- RMSEA - root mean square error of approximation is an absolute measure of fit. <= 0.05 is a close fit. 0.05 - 0.08 reasonable approximation. >-0.10 poor fit.

### 2-Factor cFA models
- 2-Factor model is poor fit in all samples
- Crossdx: CFI 0.850, TLI 0.821, RMSEA 0.095
- SSD: CFI 0.832, TLI 0.799, RMSEA 0.117
- PS: CFI 0.813, TLI 0.776, RMSEA 0.126
```{r}
#2 - factor
cmod2f <- 'f1 =~ tlc_06derail + tlc_02povcontent + tlc_05tangent + tlc_13lossgoal + tlc_08illogic + tlc_12circum + tlc_14persev + tlc_07incoh + tlc_03pressure + tlc_04distract + tlc_10neologism
f2 =~ tlc_01povspeech + sans_6_vocalinf + sans_11_latency'

#Default - test in cross-dx
cmod2f_crossdx <- cfa(cmod2f, data = crossdx.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod2f_crossdx, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 2-factor from crossdx in SSD
cmod2f_ssd <- cfa(cmod2f, data = ssd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod2f_ssd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 2-factor from crossdx in OPD
cmod2f_opd <- cfa(cmod2f, data = opd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod2f_opd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)
```

### 3-Factor cFA without cross-loadings
- 3-factor model without the cross loaded factors is very good for crossdx, borderline for SSD
- Fixing variance to 1 performs better than fixing the loading of first variable
- Performance is good for Crossdx, acceptable for SSD, and good for PS
- Crossdx: CFI 0.965, TLI 0.953, RMSEA 0.047
- SSD: CFI 0.911, TLI 0.880, RMSEA 0.085
- PS: CFI 0.945, TLI 0.927, RMSEA 0.068
```{r}
#3 - factor
cmod3f <- 'f1 =~ tlc_05tangent + tlc_12circum + tlc_03pressure + tlc_04distract + tlc_14persev
f2 =~ tlc_07incoh + tlc_08illogic + tlc_10neologism
f3 =~ tlc_01povspeech + sans_6_vocalinf + sans_11_latency'

#Default - test in cross-dx
cmod3f_crossdx <- cfa(cmod3f, data = crossdx.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod3f_crossdx, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 3-factor from crossdx in SSD
cmod3f_ssd <- cfa(cmod3f, data = ssd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod3f_ssd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 3-factor from crossdx in OPD
cmod3f_opd <- cfa(cmod3f, data = opd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod3f_opd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

```

### 4-Factor Model- with separate cross-loaded factor
- Fixing variance performs better than fixing loading of first variable
- Performance is good for crossdx, acceptable for SSD, good for psychosis spectrum group
- Crossdx: CFI 0.954, TLI 0.941, RMSEA 0.054
- SSD: CFI 0.918, TLI 0.894, RMSEA 0.085
- PS: CFI 0.940, TLI 0.923, RMSEA 0.074
```{r}
###Fourfactor model
#4 - factor
cmod4f <- 'f1 =~ tlc_05tangent + tlc_12circum + tlc_03pressure + tlc_04distract + tlc_14persev
f2 =~ tlc_07incoh + tlc_08illogic + tlc_10neologism
f3 =~ tlc_01povspeech + sans_6_vocalinf + sans_11_latency
f4 =~ tlc_06derail + tlc_13lossgoal + tlc_02povcontent'

#Cross-dx
cmod4f_crossdx <- cfa(cmod4f, data = crossdx.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod4f_crossdx, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#Test 4-factor in SSD
cmod4f_ssd <- cfa(cmod4f, data = ssd.tlc, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod4f_ssd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#OPD
cmod4f_opd <- cfa(cmod4f, data = opd.clin, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod4f_opd, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)
```

## Generate Factor Scores
- Based on eFA results
- Grice, James W.,2001, Computing and evaluating factor scores, Psychological Methods, 01,4, 430-450. (note the typo in equation 8)
- ten Berge, Jos M.F., Wim P. Krijnen, Tom Wansbeek and Alexander Shapiro (1999) Some new results on correlation-preserving factor scores prediction methods. Linear Algebra and its Applications, 289, 311-318. - Using Ten Berge method because it's able to preserve oblique-ness of factors

```{r, include=FALSE}
### 3-Factor Solution - generating factor scores
just3factortlc <- factor.scores(crossdx.tlc, model02, method = "tenBerge", impute="mean")$scores %>% as.data.frame()
str(just3factortlc)
names(just3factortlc) <- c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress")

#Get rid of any old variables from prior runs
crossdx.clin <- select(crossdx.clin, -starts_with("tlc_3f"))

# Combine with full dataset
crossdx.clin <- cbind(crossdx.clin, just3factortlc)

### filling in missing group variables 
crossdx.clin$grid[which(is.na(crossdx.clin$group))] #all good

save(crossdx.clin, file="crossdx_clin.R")
write.csv(crossdx.clin, file="crossdx_clin.csv")

### Saving Model
save(model02, file="TLC3fmodel.R")
model02$loadings

save(model01, file="TLC2fmodel.R")

```




## Plotting & Comparing Factor relationships

### Overall Group Effects
```{r, results='asis'}
tableby(group ~ tlc_3f_inefficient + tlc_3f_incoherent + tlc_3f_impexpress,
        data=crossdx.clin, numeric.stats=c("mean", "range", "sd")) %>% summary(digits = 2, digits.p=3)
```


```{r}
# Long format for plotting together
temp <- pivot_longer(crossdx.clin, cols = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), names_to = "factor")
justlongfactors <- temp %>% select(group, factor, value)
justlongfactors$factor <- factor(justlongfactors$factor, levels = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), labels = c("Inefficient Speech", "Incoherent Speech", "Impaired Expressivity"))

# Make group consistent with poster labels
justlongfactors$group <- as.character(justlongfactors$group)
justlongfactors$group[which(justlongfactors$group == "HC")] <- "HV"
justlongfactors$group[which(justlongfactors$group == "HSC")] <- "OD"
justlongfactors$group[which(justlongfactors$group == "OPD")] <- "PSY"
justlongfactors$group <- as.factor(justlongfactors$group)
table(justlongfactors$group)

#PLOT
ggplot(justlongfactors, aes(x = group, y=value, fill = group)) +
      geom_jitter(alpha=0.4) +
  geom_boxplot(alpha=0.7) + 
  facet_grid(.~factor) + 
  theme_few() + 
  scale_fill_viridis_d() + 
  labs(x = NULL, y = "Factor Z-Scores", fill = NULL) + 
  theme(text=element_text(size=20, family="Times New Roman")) +
  stat_compare_means(label = "p.signif", method = "t.test", comparisons = list( c("HV", "OD"), c("OD", "PSY"), c("PSY", "SSD"), c("HV", "PSY"), c("OD", "SSD"), c("HV", "SSD") ))

```
- multiple comparisons
```{r}
pairwise.t.test(crossdx.clin$tlc_3f_inefficient, crossdx.clin$group, p.adjust.method = "fdr")
pairwise.t.test(crossdx.clin$tlc_3f_incoherent, crossdx.clin$group, p.adjust.method = "fdr")
pairwise.t.test(crossdx.clin$tlc_3f_impexpress, crossdx.clin$group, p.adjust.method = "fdr")
```


### Group effects - matched for gender
- Used PSY as the index group because it had the fewest participants, then exact-matched from each other group.
- Using Matchit package
- optimal matching, with probit propensity score
```{r, include=FALSE}
### Matching

## Psy + HV
temp <- filter(crossdx.clin, group == "OPD" | group == "HC")
chisq.test(temp$group, temp$demo_gender)

# Checking std. Mean differences prior to Match
matchit(group ~ demo_gender, data = temp, method = NULL, distance = "glm") %>% summary()

# Match: 
mch.gender_hc <- matchit(group ~ demo_gender, data = temp, method = "optimal", distance = "glm", link="probit")
summary(mch.gender_hc)
temp2 <- match.data(mch.gender_hc)
chisq.test(temp2$group, temp2$demo_gender)

## OPD + Other disorders
temp <- filter(crossdx.clin, group == "HSC" | group == "OPD")
chisq.test(temp$group, temp$demo_gender)

# Checking std. Mean differences prior to Match
matchit(group ~ demo_gender, data = temp, method = NULL, distance = "glm") %>% summary()

# Match: 
mch.gender_hsc <- matchit(group ~ demo_gender, data = temp, method = "optimal", distance = "glm", link="probit")
summary(mch.gender_hsc)
temp1 <- match.data(mch.gender_hsc)
chisq.test(temp1$group, temp1$demo_gender)

## OPD + SSD 
temp <- filter(crossdx.clin, group == "OPD" | group == "SSD")
chisq.test(temp$group, temp$demo_gender)

# Checking std. Mean differences prior to Match
matchit(group ~ demo_gender, data = temp, method = NULL, distance = "glm") %>% summary()

# Match:
mch.gender_ssd <- matchit(group ~ demo_gender, data = temp, method = "optimal", distance = "glm", link="probit")
summary(mch.gender_ssd)
temp3 <- match.data(mch.gender_ssd)
chisq.test(temp3$group, temp3$demo_gender)

### Compile
mch.gender <- temp2
temp1 <- filter(temp1, group == "HSC")
mch.gender <- rbind(mch.gender, temp1)
temp3 <- filter(temp3, group == "SSD")
mch.gender <- rbind(mch.gender, temp3)

table(mch.gender$group, mch.gender$demo_gender)
```
- Demo of matched sample
```{r, results='asis'}
tableby(group ~ demo_age + demo_gender + demo_race + demo_ethnicity + tlc_global,
        data=mch.gender, numeric.stats=c("mean", "range", "sd")) %>% summary(digits = 2, digits.p=3)
```

- Examine group effects in gender-matched sample: consistent!
```{r}
# Long format for plotting together
temp <- pivot_longer(mch.gender, cols = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), names_to = "factor")
mch.gender.longfactors <- temp %>% select(group, factor, value)
mch.gender.longfactors$factor <- factor(mch.gender.longfactors$factor, levels = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), labels = c("Inefficient Speech", "Incoherent Speech", "Impaired Expressivity"))

# Make group consistent with poster labels
mch.gender.longfactors$group <- as.character(mch.gender.longfactors$group)
mch.gender.longfactors$group[which(mch.gender.longfactors$group == "HC")] <- "HV"
mch.gender.longfactors$group[which(mch.gender.longfactors$group == "HSC")] <- "OD"
mch.gender.longfactors$group[which(mch.gender.longfactors$group == "OPD")] <- "PSY"
mch.gender.longfactors$group <- as.factor(mch.gender.longfactors$group)
table(mch.gender.longfactors$group)

#PLOT
ggplot(mch.gender.longfactors, aes(x = group, y=value, fill = group)) +
      geom_jitter(alpha=0.4) +
  geom_boxplot(alpha=0.7) + 
  facet_grid(.~factor) + 
  theme_few() + 
  scale_fill_viridis_d() + 
  labs(x = NULL, y = "Factor Z-Scores", fill = NULL) + 
  theme(text=element_text(size=20, family="Times New Roman")) +
  stat_compare_means(label = "p.signif", method = "t.test", comparisons = list( c("HV", "OD"), c("OD", "PSY"), c("PSY", "SSD"), c("HV", "PSY"), c("OD", "SSD"), c("HV", "SSD") ))

```


### Group effects - matched for age + Race
- Dropped YT sample because no age/race, only 19 OPD, so dropped this as well
- Looked at white/black race because these were the largest groups
- Used HV as the index group because it had the fewest participants, then matched from each other group.
- Using Matchit package - optimal matching, with probit propensity score
```{r, include=FALSE}
### setting up
justnoyt <- filter(crossdx.clin, study!="YT")
table(justnoyt$group)
table(justnoyt$demo_race)

# White/Black Race 
justnoyt$demo_white <- rep(0, dim(justnoyt)[1])
justnoyt$demo_black <- rep(0, dim(justnoyt)[1])

justnoyt$demo_white[which(justnoyt$demo_race == "White")] <- 1
justnoyt$demo_black[which(justnoyt$demo_race == "Black")] <- 1

table(justnoyt$demo_white)
table(justnoyt$demo_black)
      
### Matching

## HC + HSC
temp <- filter(justnoyt, group == "HSC" | group == "HC")
chisq.test(temp$group, temp$demo_white)
chisq.test(temp$group, temp$demo_black)
t.test(temp$demo_age ~ temp$group)

# Checking std. Mean differences prior to Match
matchit(group ~ demo_age + demo_white + demo_black, data = temp, method = NULL, distance = "glm") %>% summary()

# Match: 
mch.age_hsc <- matchit(group ~ demo_age + demo_white + demo_black, data = temp, method = "optimal", distance = "glm", link="probit")
summary(mch.age_hsc)
temp1 <- match.data(mch.age_hsc)
chisq.test(temp1$group, temp1$demo_race)
t.test(temp1$demo_age ~ temp1$group)

## HC + SSD 
temp <- filter(justnoyt, group == "SSD" | group == "HC")
chisq.test(temp$group, temp$demo_white)
chisq.test(temp$group, temp$demo_black)
t.test(temp$demo_age ~ temp$group)

# Checking std. Mean differences prior to Match
matchit(group ~ demo_age + demo_white + demo_black, data = temp, method = NULL, distance = "glm") %>% summary()

# Match: 
mch.age_ssd <- matchit(group ~ demo_age + demo_white + demo_black, data = temp, method = "optimal", distance = "glm", link="probit")
summary(mch.age_ssd)
temp2 <- match.data(mch.age_ssd)
chisq.test(temp2$group, temp2$demo_race)
t.test(temp2$demo_age ~ temp2$group)


### Compile
mch.age <- temp1
temp2 <- filter(temp2, group == "SSD")
mch.age <- rbind(mch.age, temp2)

```
- age + race matched group statistics
```{r, results='asis'}
tableby(group ~ demo_age + demo_gender + demo_race + demo_ethnicity + tlc_global,
        data=mch.age, numeric.stats=c("mean", "range", "sd")) %>% summary(digits = 2, digits.p=3)

tableby(group ~ demo_age + demo_gender + demo_race + demo_ethnicity + tlc_global,
        data=mch.age, numeric.stats=c("median")) %>% summary(digits = 2, digits.p=3)
```
- Examine group effects in age/race-matched sample: consistent!
```{r}
# Long format for plotting together
temp <- pivot_longer(mch.age, cols = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), names_to = "factor")
mch.age.longfactors <- temp %>% select(group, factor, value)
mch.age.longfactors$factor <- factor(mch.age.longfactors$factor, levels = c("tlc_3f_inefficient", "tlc_3f_incoherent", "tlc_3f_impexpress"), labels = c("Inefficient Speech", "Incoherent Speech", "Impaired Expressivity"))

# Make group consistent with poster labels
mch.age.longfactors$group <- as.character(mch.age.longfactors$group)
mch.age.longfactors$group[which(mch.age.longfactors$group == "HC")] <- "HV"
mch.age.longfactors$group[which(mch.age.longfactors$group == "HSC")] <- "OD"
mch.age.longfactors$group <- as.factor(mch.age.longfactors$group)
table(mch.age.longfactors$group)

#PLOT
ggplot(mch.age.longfactors, aes(x = group, y=value, fill = group)) +
      geom_jitter(alpha=0.4) +
  geom_boxplot(alpha=0.7) + 
  facet_grid(.~factor) + 
  theme_few() + 
  scale_fill_viridis_d() + 
  labs(x = NULL, y = "Factor Z-Scores", fill = NULL) + 
  theme(text=element_text(size=20, family="Times New Roman")) +
  stat_compare_means(label = "p.signif", method = "t.test", comparisons = list( c("HV", "OD"), c("OD", "SSD"), c("HV", "SSD") ))

```

## Factor Solution and Data Collection method
- YT vs. others since recording method should not impact TLC ratings (same task)
- YT 3 Factor: CFI 0.940, TLI 0.916, RMSEA 0.091
- YT 4 Factor: CFI 0.888, TLI 0.852, RMSEA 0.127
- No-YT 3 Factor: CFI 0.949, TLI 0.932, RMSEA 0.056
- No-YT 4 factor: CFI 0.946, TLI 0.930, RMSEA 0.058
```{r}
### Just YT
justyt_cfa <- filter(crossdx.clin, study=="YT")
justyt_cfa <- select(justyt_cfa, all_of(vars_fa))
summary(justyt_cfa)

#3f: drop distractibility because all 0's
cmod3f_notlc4 <- 'f1 =~ tlc_05tangent + tlc_12circum + tlc_03pressure + tlc_14persev
f2 =~ tlc_07incoh + tlc_08illogic + tlc_10neologism
f3 =~ tlc_01povspeech + sans_6_vocalinf + sans_11_latency'

#3-Factor CFA
cmod3f_yt <- cfa(cmod3f_notlc4, data = justyt_cfa, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod3f_yt, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#4F: drop distractibility
cmod4f_notlc4 <- 'f1 =~ tlc_05tangent + tlc_12circum + tlc_03pressure + tlc_14persev
f2 =~ tlc_07incoh + tlc_08illogic + tlc_10neologism
f3 =~ tlc_01povspeech + sans_6_vocalinf + sans_11_latency
f4 =~ tlc_06derail + tlc_13lossgoal + tlc_02povcontent'

#4factor CFA
cmod4f_yt <- cfa(cmod4f_notlc4, data = justyt_cfa, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod4f_yt, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)


### No YT
justnoyt_cfa <- filter(crossdx.clin, study!="YT")
justnoyt_cfa <- select(justnoyt_cfa, all_of(vars_fa))

#3-Factor CFA
cmod3f_noyt <- cfa(cmod3f, data = justnoyt_cfa, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod3f_noyt, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)

#4factor CFA
cmod4f_noyt <- cfa(cmod4f, data = justnoyt_cfa, std.lv = TRUE, estimator = "MLR", missing = "ml")

summary(cmod4f_noyt, fit.measures = TRUE, standardized = TRUE, modindices = FALSE, rsquare=TRUE)


```

